---
title: "Attainable fourth order concordance"
author: "Alexander J. McNeil and Johanna Neslehova"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Attainable fourth order concordance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(extremalcopula)
library(limSolve)
```

In this example we take a 4 dimensional copula and assume that all the pairwise probabilities of concordance are equal to some value $\kappa_2$. We determine first the minimum value for the 4th order probability of concordance $\kappa_4$ and then the maximum value using linear programming. These can be checked against the true values which can be determined analytically without recourse to optimization.

```{r, echo=TRUE}
kappa2 <- 0.7
partsignature <- c(1, rep(kappa2, 6))

attainable(partsignature)
max(2*kappa2 -1, 0)

attainable(partsignature, maximize = TRUE)
(3*kappa2 -1)/2
```


