---
title: "Generating mixtures of extremal copulas"
author: "Alexander J. McNeil and Johanna Neslehova"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating mixtures of extremal copulas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(extremalcopula)
```
### Generating mixtures of extremal copulas
First we illustrate how each extremal copula is coded with a binary number.
```{r, echo=TRUE}
binarycoding(c(1:8), 4)
```

Suppose we wanted to generate 10 realisations from copula 2 in 4 dimensions.
```{r, echo=TRUE}
U <- rextremal(rep(2, 10), 4)
U
cor(U)
```

Now suppose wanted generate a mixture from all 4 dimensional copulas. Here we use an increasing series of weights. The Pearson and Kendall correlations matrice of the extremal mixture copula are identical and their sample estimates should be close.
```{r, echo=TRUE}
set.seed(17)
U <- rextremalmix((1:8)/36, 1000)
cor(U)
cor(U, method = "kendall")
```


