---
title: "A Kendall matrix that is not elliptical"
author: "Alexander J. McNeil and Johanna Neslehova"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Kendall matrix that is not elliptical}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(extremalcopula)
library(limSolve)
```


Here is a $4 \times 4$ correlation matrix. Note that it is positive definite.

```{r, echo=TRUE}
P <- copula::p2P(c(-0.19, -0.29, 0.49, -0.34, 0.3, -0.79))
P

eigen(P)$values
```

Could it be a matrix of Kendall's tau values? On the assumption that it is, let's calculate the pairwise probabilities of concordance. More precisely we find the first 7 elements of the concordance signature.

```{r, echo=TRUE}
kappa <- c(1, (1 + P[lower.tri(P)])/2)
kappa
```

Let's see if we can find an extremal mixture copula that matches these probabilities of concordance. 

```{r, echo=TRUE}
weights <- attainable(kappa)
weights
```

This vector of weights seems to do it. We can check by computing the weighted sum of the extremal correlation matrices.

```{r, echo=TRUE}
Reduce('+', Map('*', weights, extcorr(1:8, 4)))
```

Here is another set of weights that would also do it.
```{r, echo=TRUE}
weights2 <- attainable(kappa, maximize = TRUE)
weights2
Reduce('+', Map('*', weights2, extcorr(1:8, 4)))
```

Can we get this set of Kendall's tau values from an elliptical copula? Let's calculate the implied correlation matrix of the copula and check it is positive semi-definite.
```{r, echo=TRUE}
P2 <- sin(pi * P / 2)
eigen(P2)$values
```

It isn't! So $P$ is an example of a Kendall's tau matrix that cannot be associated with an elliptically distributed random vector.
